---
title: "lab5"
author: "Sai Gugulothu"
date: "4/18/2019"
output: html_document
---

```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
names(acc)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
names(acc)

```

```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyverse)
```
Part1: Accident Data 
Question: are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?

I used ggplot to answer the question and you can see that most accidents happen on Fridays and the weekend. 

```{r}
ggplot(data = acc, aes(x=DAY_WEEK)) + geom_histogram(color="black", fill="yellow") + ylab("Total Accidents") + xlab("Day of the Week") + scale_x_continuous(breaks=c(1, 2, 3, 4, 5, 6, 7),labels=c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))+ggtitle("Accidents that Occur on Different Days of Week")




```

Question: What time of the day do accidents happen (use variable HOUR)?

For this question I also used ggplot to plot a bar chart. You can see form the graph that there are more accident in the evening hours from around 6PM-7PM ish.  

```{r}
ggplot(acc, aes(x=factor(HOUR))) + geom_bar(colour = "black", fill = "blue") + xlim(c(0, 23))  + scale_x_discrete(name ="Time of day (hours)", limits=1:23)

```


Question: What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?

There are 8769 accidents with at least one drunk driver. 


```{r}
sum(acc$DRUNK_DR >= 1)
```

Part 2: Connecting data

```{r}
drive <- per %>% filter(PER_TYP==1)
joined <- inner_join(acc, drive, by="ST_CASE")
joined$SEX[joined$SEX %in% c(8,9)] <- NA
joined$SEX <- factor(joined$SEX)
levels(joined$SEX) <- c("MALE", "FEMALE")
joined <- filter(joined)
ggplot(joined, aes(x=HOUR.x, fill=SEX)) + geom_bar()  + xlab("Hours") + ylab("Number of People") + ggtitle("Accidents Grouped by Stats") + facet_grid(DAY_WEEK~.)


```


Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!


Part 3: Connecting data and make maps 

```{r}
#us <- readxl::read_xlsx("FRPP GLC United States.xlsx")
#territories <- readxl::read_xlsx("FRPP GLC US Territories.xlsx")


```



