---
title: "individual2.Rmd"
author: "Sonya Haan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 5

# Part 1: Accident Data

**Load the data and packages**
```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
names(acc)
person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
names(person)

library(ggplot2)
library(tidyverse)
```

**Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?**

```{r}
acc$DAY_WEEK <- factor(acc$DAY_WEEK)
levels(acc$DAY_WEEK) <- c("SUN", "MON", "TUE", "WED", "THU", "FRI", "SAT")
table(acc$DAY_WEEK)
ggplot(acc, aes(x=DAY_WEEK)) + geom_bar() + ggtitle("Accidents per day of week")
```

According to the data, more accidents occur on weekends than on weekdays. The day with the most number of accidents is Saturday.

**What time of the day do accidents happen (use variable HOUR)?**

```{r}
acc$HOUR[acc$HOUR==99] <- is.na(acc$HOUR)
acc$HOUR <- factor(acc$HOUR)
table(acc$HOUR)
ggplot(acc, aes(x=HOUR)) + geom_bar() + ggtitle("Accidents per time of day")
```

Most accidents occur between 2pm and midnight. The time that the highest number of accidents occurred is 6pm.

**What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?**

```{r}
drunk <- acc %>% filter(DRUNK_DR >= 1)
count(drunk)
```

The number of accidents involving one or more drunk drivers is 8,769.

# Part 2: Connecting Data

**Load the person table. Identify drivers (PER_TYP == 1, see fars manual) and subset on them.**

```{r}
drivers <- person %>% filter(PER_TYP==1)
```

**Join accident and driver table (work out which variable(s) to use)**

```{r}
accAndDrivers <- inner_join(acc, drivers, by="ST_CASE")
```

**Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!**

```{r}
accAndDrivers$SEX[accAndDrivers$SEX %in% c(8,9)] <- NA
accAndDrivers$SEX <- factor(accAndDrivers$SEX)
levels(accAndDrivers$SEX) <- c("MALE", "FEMALE")
accAndDrivers <- filter(accAndDrivers, !is.na(SEX))
ggplot(accAndDrivers, aes(x=HOUR.x, fill=SEX)) + geom_bar(na.rm=TRUE) + facet_grid(DAY_WEEK~.) + ggtitle("Accidents per day, hour, and gender")
```

# Part 3: Connecting Data and Make Maps

**Visualize the number of accidents in each county using a choropleth map. To do this, you need to join the accident data, GLCs, and appropriate map data. Can you find seasonal trends in accident occurrences in different states? Is summer or winter more dangerous? Why?**

```{r}
glc <- readxl::read_xlsx("FRPP GLC United States.xlsx")
colnames(glc) = glc[1, ]
glc = glc[-1, ]  
glc$`County Code` <- as.numeric(glc$`County Code`)
glc$`State Code` <- as.numeric(glc$`State Code`)
glc <- glc %>% select(`State Name`, `State Code`, `County Name`, `County Code`)
glc <- unique(glc)
accGlc <- left_join(acc, glc, by=c("STATE"="State Code", "COUNTY"="County Code")) %>% select("STATE", "ST_CASE", "COUNTY", "MONTH", "State Name", "County Name")
accGlc <- accGlc %>% mutate("STATE_NAME"=tolower(accGlc$`State Name`), "COUNTY_NAME"=tolower(accGlc$`County Name`))
accGlc <- accGlc %>% select(-"State Name", -"County Name")
accGlcState <- accGlc
accGlc <- accGlc %>% group_by(STATE_NAME, COUNTY_NAME) %>% mutate("Count"=n()) %>% select(-"ST_CASE")
accGlc <- unique(accGlc)

counties <- map_data("county")

accGlc.map <- accGlc %>% left_join(counties, by=c("STATE_NAME"="region", "COUNTY_NAME"="subregion"))

accGlc.map %>%
  ggplot(aes(x = long, y = lat, fill=Count)) +
  geom_polygon(aes(group=group)) + ggtitle("Accidents by County")

accGlcState <- accGlc %>% group_by(STATE_NAME, MONTH) %>% mutate("Count"=n()) %>% select("MONTH","STATE_NAME","Count")
accGlcState <- unique(accGlcState)

states <- map_data("state")
accGlcState.map <- accGlcState %>% left_join(states, by=c("STATE_NAME"="region"))
accGlcState.map %>% filter(MONTH==6) %>%
  ggplot(aes(x = long, y = lat, fill=Count)) +
  geom_polygon(aes(group=group)) + ggtitle ("Accidents per State (June)")

accGlcState.map %>% filter(MONTH==12) %>%
  ggplot(aes(x = long, y = lat, fill=Count)) +
  geom_polygon(aes(group=group)) + ggtitle ("Accidents per State (December)")
```

Overall it appears that slightly more accidents occur in a summer month like June compared to a winter month like December (note the scale goes to 100 in June but only 90 in December). That being said, states have about the same number of accidents comparative to other states regardless of the season. (Texas has many more accidents than other states regardless of the month.)





#Sai Gugulothu

---
title: "lab5"
author: "Sai Gugulothu"
date: "4/18/2019"
output: html_document
---

```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
names(acc)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
names(acc)

```

```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyverse)
```
Part1: Accident Data 
Question: are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?

I used ggplot to answer the question and you can see that most accidents happen on Fridays and the weekend. 

```{r}
ggplot(data = acc, aes(x=DAY_WEEK)) + geom_histogram(color="black", fill="yellow") + ylab("Total Accidents") + xlab("Day of the Week") + scale_x_continuous(breaks=c(1, 2, 3, 4, 5, 6, 7),labels=c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))+ggtitle("Accidents that Occur on Different Days of Week")




```

Question: What time of the day do accidents happen (use variable HOUR)?

For this question I also used ggplot to plot a bar chart. You can see form the graph that there are more accident in the evening hours from around 6PM-7PM ish.  

```{r}
ggplot(acc, aes(x=factor(HOUR))) + geom_bar(colour = "black", fill = "blue") + xlim(c(0, 23))  + scale_x_discrete(name ="Time of day (hours)", limits=1:23)

```


Question: What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?

There are 8769 accidents with at least one drunk driver. 


```{r}
sum(acc$DRUNK_DR >= 1)
```

Part 2: Connecting data

```{r}
drive <- per %>% filter(PER_TYP==1)
joined <- inner_join(acc, drive, by="ST_CASE")
joined$SEX[joined$SEX %in% c(8,9)] <- NA
joined$SEX <- factor(joined$SEX)
levels(joined$SEX) <- c("MALE", "FEMALE")
joined <- filter(joined)
ggplot(joined, aes(x=HOUR.x, fill=SEX)) + geom_bar()  + xlab("Hours") + ylab("Number of People") + ggtitle("Accidents Grouped by Stats") + facet_grid(DAY_WEEK~.)


```


Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!


Part 3: Connecting data and make maps 

```{r}
#us <- readxl::read_xlsx("FRPP GLC United States.xlsx")
#territories <- readxl::read_xlsx("FRPP GLC US Territories.xlsx")


```



